---
title: "Emoji Sentiment Analysis"
author: "Ryan Wesslen"
date: "July 20, 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Emoji Analysis 

This analysis is based on a [blog post](http://opiateforthemass.es/articles/emoji-analysis/) and [code](https://github.com/today-is-a-good-day/emojis/blob/master/emoji_analysis.R) by Jessica Peterka-Bonetta.

I only made small customizations in the code.

### Load the dataset

First, let's reload our dataset.

```{r warning=FALSE}
#install.packages(streamR)
library(streamR)

file <- "../data/stream/stream_got.json"
# file <- "~/Dropbox (UNC Charlotte)/summer-2017-social-media-workshop/data/stream/stream_got.json"

#?parseTweets
tweets <- parseTweets(tweets = file)

tweets$text <- iconv(tweets$text, from="UTF-8", to="ASCII", "byte")
```

Next, let's create a dictionary for all of the emojis.

```{r warning=FALSE}
library(tidyverse); library(stringr); library(Unicode)

source("./emoji_functions.R")
emoji_file = "../data/emojis.csv"

emDict <- readr::read_delim(file = emoji_file, delim = ";")

emDict <- emDict %>%
  mutate(description = tolower(EN)) %>%
  mutate(unicode = as.u_char(unicode))
```

Then, let's match the emojis to count the most used emojis.

```{r}
#helper functions
matchto <- emDict$ftu8
description <- emDict$EN

rank <- emojis_matching(tweets$text, matchto, description) %>% 
  group_by(description) %>% 
  summarise(n = sum(count)) %>%
  arrange(-n)

head(rank, n = 10)
```

### Emoji Sentiment Analysis

We can then do an emoji "sentiment analysis".

To do this, we need [this site](http://kt.ijs.si/data/Emoji_sentiment_ranking/index.html) to provide sentiment scores of the tweets.

```{r warning=FALSE}
library(xml2); library(rvest)

# reference website
url <- "http://kt.ijs.si/data/Emoji_sentiment_ranking/index.html"

# get emoticons
emojis_raw <- url %>%
  read_html() %>%
  html_table() %>%
  data.frame %>%
  select(-Image.twemoji., -Sentiment.bar.c.i..95..)

# rename the file
names(emojis_raw) <- c("char", "unicode", "occurrences", "position", "negative", "neutral", 
                   "positive", "sentiment_score", "description", "block")

# change numeric unicode to character unicode to be able to match with emDict 
emojis <- emojis_raw %>%
  mutate(unicode = as.u_char(unicode)) %>%
  mutate(description = tolower(description)) 

# merge with emDict to get encoding
emojis_merged <- emojis %>%
  merge(emDict, by = "unicode")

new_matchto <- emojis_merged$ftu8
new_description <- emojis_merged$description.x
sentiment <- emojis_merged$sentiment_score

sentiments <- emojis_matching(tweets$text, new_matchto, new_description, sentiment) %>%
  mutate(sentiment = count*as.numeric(sentiment)) %>%
  group_by(text) %>% 
  summarise(sentiment_score = sum(sentiment), tweet_count = n()) %>%
  filter(!is.na(sentiment_score)) %>%
  arrange(desc(sentiment_score))

# top 10 most positive tweets by emoji sentiment
head(sentiments, n = 10)
```

