---
title: "Twitter Social Networks"
author: "Ryan Wesslen"
date: July 20, 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Retweet Network

### Loading JSON file

```{r warning=FALSE}
#install.packages(streamR)
library(streamR); library(tidyverse)

file <- "../data/stream/stream_got.json"
#file <- "~/Dropbox (UNC Charlotte)/summer-2017-social-media-workshop/data/stream/stream_got.json"

#?parseTweets
tweets <- parseTweets(tweets = file)
```

### Creating a Retweet Network

In this example, we'll create a retweet network. This snippet of code was from [Pablo Barbera](https://github.com/pablobarbera/big-data-upf/blob/master/02d-twitter-streaming-data-collection.Rmd), who created the `streamR` package and who has phenomenal R materials for analyzing Twitter (highly recommend his GitHub!).

```{r}
library(igraph)

# regular expression -- get only tweets that have "RT @"
rts <- tweets[grep("RT @", tweets$text),]

edges <- data.frame(
  node1 = rts$screen_name,
  node2 = gsub('.*RT @([a-zA-Z0-9_]+):? ?.*', rts$text, repl="\\1"),
  stringsAsFactors=F
)

g <- graph_from_data_frame(d=edges, directed=TRUE)
```

Now we have a network as a `igraph` object g.

Let's run some basic network stats, e.g. find total number of connections.

```{r}
degTotal <- degree(g, mode = "total")
degTotal <- degTotal[order(degTotal, decreasing = T)]

head(degTotal, n = 10)
```

Alternatively, since this is a directed network, we can find who did the most retweets (i.e., retweeted others).

```{r}
degOut <- degree(g, mode = "out")
degOut <- degOut[order(degOut, decreasing = T)]

head(degOut, n = 10)
```

Remember -- our dataset was only for 10 mins. So those people did a lot of #gameofthrones retweets in only 10 minutes.

Further, we can find who was retweeted the most.

```{r}
degIn <- degree(g, mode = "in")
degIn <- degIn[order(degIn, decreasing = T)]

head(degIn, n = 10)
```

Let's instead observe the network using `visNetwork`, a [HTMLWidget](http://gallery.htmlwidgets.org)

```{r}
library(visNetwork)

V(g)$size <- 10

visIgraph(g) %>% 
  visInteraction(navigationButtons = TRUE) 
```

However, things are a bit "squished" together. An important property of social networks is to alter the layout.

For example, we can set a property to add in "physics".

```{r}
visIgraph(g) %>% 
  visInteraction(navigationButtons = TRUE) %>%
  visIgraphLayout("physics" = TRUE)
```

We can also change the size of the nodes to correspond to the number of (total) degrees.

```{r}
V(g)$size <- log(degTotal) +1

visIgraph(g) %>% 
  visInteraction(navigationButtons = TRUE) %>%
  visIgraphLayout("physics" = TRUE)
```

Just for fun, we can use another HTMLWidget [edgebundler](https://github.com/garthtarr/edgebundleR) to create a different type of network.

```{r results='asis'}
#install.packages("edgebundleR")
library(edgebundleR)

edgebundle(g, fontsize = 6, directed = TRUE)
```

There are a TON of other ways to customize networks. I **strongly** recommend users who are interested in analyzing social networks to check out Katya Ognyanova's [phenomenal network tutorial](http://kateto.net/network-visualization) for a wider range of options.

## Challenge: Creating other Network

### Reply Network (Easy)

For this challenge, you'll need to create a reply network.

For this, you'll need to use similar code in rows 111-117. However, instead of focusing on the RT's, you'll simply need to create the edge list from two fields: `screen_name` (from) and `in_reply_to_screen_name`.

Like the RT example, you need to first create a subset of the tweets to only include tweets that include a reply. You can use the `subset` function (type ?subset to look it up) along with the is.na() function to identify only tweets that have a reply. Make sure to take the "negation" (!) of the is.na function.

See the attached R code (challenge-solution.R) for the solution.

```{r}
## write your response here

```

### Mention Networks (Hard/just FYI)

If you're able to complete the first task, try to run this code for a mention network

Code from this section is from cosmopolitanvan's [RPubs tutorial](https://rpubs.com/cosmopolitanvan/twitternetworks), which is another great resource.

```{r warning=FALSE}
# regular expression -- get only tweets that have "@" (i.e., mention)
mention <- tweets[grep("@", tweets$text),]

names <- lapply(mention$text, function(tx) {
    matches = gregexpr('@[^([:blank:]|[:punct:])]+', tx)[[1]]
    sapply(seq_along(matches), function(i) 
      substr(tx, matches[i] + 1, matches[i] + attr(matches, 'match.length')[i] - 1))
  })

edges3 = 
  lapply(seq_along(mention$text), function(i) {
    if(names[[i]] == '')  
      return(NULL)
    lapply(names[[i]], function(m)
      c(node1 = as.character(mention$screen_name[i]), node2 = m)) %>%
      do.call(rbind, .) %>% as.data.frame()
  }) %>% 
  do.call(rbind, .) %>%
  count(tolower(node1), tolower(node2))

g3 <- graph_from_data_frame(d = edges3, directed = T)

V(g3)$size <- 10

visIgraph(g3) %>% 
  visInteraction(navigationButtons = TRUE) %>%
  visIgraphLayout("physics" = TRUE)
```